Theories of deliberated choice

### 2022-11-01 simple.tex
Attempt more precise Tf

While $I$ and $\allargs$ are fundamental sets whose elements are considered elementary, a few notations and concepts are required to define ${\allleadsto}$.
Throughout the article, $\N$ includes zero. %and $\N^* = \N \setminus \set{0}$. 
Given $j, l \in \N$, the notation $\intvl{j, l}$ represents $\set{k \in \N \suchthat j ≤ k ≤ l}$ (thus $\forall j \in \N: \intvl{j, 0} = \emptyset$).
Define $\allhist = \bigcup_{k \in \N} \allargs^{\intvl{1, k}}$ as the finite sequences of arguments, including the empty sequence. 
A generic element of $\allhist$ (a sequence of arguments) is denoted by $\hist$ and called an argumentative path, or simply a path. Its length is denoted by $\card{\hist} \in \N$.
Thus, given an argumentative path $\hist \in \allhist$, $\forall k \in \intvl{1, \card{\hist}}: \hist_k \in \allargs$, and $\hist = \emptyset ⇔ \card{\hist} = 0$.

@@

An argumentative theory claims that whenever some of its arguments fail to convince $i$ of choosing $\pga$, it can argue something else.
It is said to be truthful iff this claim is correct.
The following definition makes this precise.%
\footnote{To be fixed.
If $a_1$, used by the theory, is not decisive, then the theory has an answer, namely, $a_2$, which works at least temporarily.}
\begin{definition}[Truthfulness]
	\label{def:truth}
	An argumentative theory $(\pga, \arga, \dg)$ for the exposure protocol $(I, \allargs, {\leadsto})$ is \emph{truthful} iff $\forall i \in I, \hist \in \allhist, \ar \in \histend, \ar_1 \in \histend: [\hist \nileadsto \pga] ⇒ [\dg(\ar_1, \hist) ≠ \emptyset \land (\hist, \dg(\ar_1, \hist)) \ileadsto \pga]$.
\end{definition}
\begin{axiom}[Answerability]
	\label{ax:ans}
	$\forall \ar, \ar': [\exists \hist \suchthat (\hist, \ar, \ar') \nileadsto \pga \land \exists \hist \suchthat (\hist, \ar, \ar') \ileadsto \pga] ⇒ \exists \hist, \ar_c \suchthat (\hist, \ar, \ar', \ar_c) \nileadsto \pga$.
\end{axiom}
\footnote{$a_1, a_2, a_3$ NOT and $a_1, a_2, a_3, a_4$ YES and $a_1', a_2, a_3, a_4$ NO imply that $a_1, a_2, a_3, a_4, a'_1$ NO.
Generalizing: $\alpha$ NO and $\alpha, a_4$ YES and $\alpha', a_4$ NO imply that $\exists a' \in \alpha' \suchthat \alpha, a_4, a'$ NO.
Let’s try with this too powerful axiom, but then we should weaken it, as it could be that $a_1$ AND $a_4$ TOGETHER are decisive, whence the axiom fails. We should have then some argument $a_1 + a_4$ to replace them, and…?}
\begin{axiom}[Reinstatement]
	\label{ax:r}
	$\forall \hist, \ar_1 \in \gargsi: [(\hist, a_1, a) \nileadsto p \land (\hist, a_1, a, a_2, …) \ileadsto p] ⇒ [(\hist, a_2) \ileadsto p \land \forall \hist': (\hist', a_2, a') \ileadsto p ⇒ (\hist', a_1, a') \nileadsto p]$.
\end{axiom}
\begin{example}[No reinstatement]
	Reinstatement is required. Consider $a_1$ not decisive because of $a$ and then $(a_1, a, a_2) \ileadsto p$. But still $a_2$ not decisive: $a2 \nileadsto p$.
\end{example}
\begin{theorem}[Validity]
	If an argumentative theory $(\pga, \arga, \dg)$ is truthful and $\allargs$ is finite, then \cref{ax:ans,ax:r} imply that the theory is correct.
\end{theorem}
\begin{proof}[Sketch of proof]
	Consider any $i$. We have to show that $\pga \in \Phi_i$.%
	\footnote{Assume $(a_4, \arga, a) \nileadsto \pga$ (thus, $a$ attacks $\arga$). 
	By tf, $(a_4, \arga, a, \ar_2, …) \ileadsto \pga$.
	We get $(a_3, \arga, a, \ar_2) \ileadsto \pga$ by answ.
	By reinst: $(\ar_2)$ YES.}

>>>>

While $I$ and $\allargs$ are fundamental sets whose elements are considered elementary, a few notations and concepts are required to define ${\allleadsto}$.
Throughout the article, $\N$ includes zero. %and $\N^* = \N \setminus \set{0}$. 
Given $j, l \in \N$, the notation $\intvl{j, l}$ represents $\set{k \in \N \suchthat j ≤ k ≤ l}$ (thus $\forall j \in \N: \intvl{j, 0} = \emptyset$).
Define $\allhist = \bigcup_{k \in \N} \allargs^{\intvl{1, k}}$ as the finite sequences of arguments, including the empty sequence. 
A generic element of $\allhist$ (a sequence of arguments) is denoted by $\hist$ and called an argumentative path, or simply a path. Its length is denoted by $\card{\hist} \in \N$.
Thus, given an argumentative path $\hist \in \allhist$, $\forall k \in \intvl{1, \card{\hist}}: \hist(k) = \hist_k \in \allargs$, and $\hist = \emptyset ⇔ \card{\hist} = 0$.

@@

An argumentative theory claims that whenever some of its arguments fail to convince $i$ of choosing $\pga$, it can argue something else.
It is said to be truthful iff this claim is correct.
The following definition makes this precise.%
\footnote{To be fixed.
If $a_1$, used by the theory, is not decisive, then the theory has an answer, namely, $a_2$, which works at least temporarily.
If $a_1$ used by th is not dec, and $a_2$ used by th is not dec *in the same sequence*, then $a_3$ needs to defend *in the same sequence*.
$(…, a_1)$ NO and $(…, a_1, …, a_2)$ NO then $a_3$ defends it.
$\alpha$ of the form $(\alpha^1, \alpha^2)$ where $a_1$ in end of $α^1$ NO and $a_2$ in end of $α^2$ NO.}
In the following definition, $(\hist^1, …, \hist^j, …, \hist^n)$ represent the sequence $\hist^1$ to which is suffixed the sequence $\hist^2$ and so on.
\begin{definition}[Truthfulness]
	\label{def:truth}
	An argumentative theory $(\pga, \arga, \dg)$ for the exposure protocol $(I, \allargs, {\leadsto})$ is \emph{truthful} iff 
	$\forall i \in I, \hist \in \allhist \suchthat 
	k_j \in \intvl{1, \card \hist} \land
	\hist^j = (\hist_1, …, \hist_{k_j}) \land
	\forall j: [k_{j - 1} < k_j \land \hist^j \nileadsto \pga \land \ar_1 = \arga \land \ar_j \in \histend[j] \land \ar_{j + 1} = \dg(\ar_j, \hist^j)]: (\hist, \ar_{n + 1}) \ileadsto \pga]$.
\end{definition}
\begin{axiom}[Answerability]
	\label{ax:ans}
	$\forall \ar, \ar': [\exists \hist \suchthat (\hist, \ar, \ar') \nileadsto \pga \land \exists \hist \suchthat (\hist, \ar, \ar') \ileadsto \pga] ⇒ \exists \hist, \ar_c \suchthat (\hist, \ar, \ar', \ar_c) \nileadsto \pga$.
\end{axiom}
\footnote{$a_1, a_2, a_3$ NOT and $a_1, a_2, a_3, a_4$ YES and $a_1', a_2, a_3, a_4$ NO imply that $a_1, a_2, a_3, a_4, a'_1$ NO.
Generalizing: $\alpha$ NO and $\alpha, a_4$ YES and $\alpha', a_4$ NO imply that $\exists a' \in \alpha' \suchthat \alpha, a_4, a'$ NO.
Let’s try with this too powerful axiom, but then we should weaken it, as it could be that $a_1$ AND $a_4$ TOGETHER are decisive, whence the axiom fails. We should have then some argument $a_1 + a_4$ to replace them, and…?}
\begin{axiom}[Reinstatement]
	\label{ax:r}
	$\forall \hist, \ar_1 \in \gargsi: [(\hist, a_1, a) \nileadsto p \land (\hist, a_1, a, a_2, …) \ileadsto p] ⇒ [(\hist, a_2) \ileadsto p \land \forall \hist': (\hist', a_2, a') \ileadsto p ⇒ (\hist', a_1, a') \nileadsto p]$.
\end{axiom}
\begin{example}[No reinstatement]
	Reinstatement is required. Consider $a_1$ not decisive because of $a$ and then $(a_1, a, a_2) \ileadsto p$. But still $a_2$ not decisive: $a2 \nileadsto p$.
\end{example}
\begin{theorem}[Validity]
	If an argumentative theory $(\pga, \arga, \dg)$ is truthful and $\allargs$ is finite, then \cref{ax:ans,ax:r} imply that the theory is correct.
\end{theorem}
\begin{proof}[Sketch of proof]
	Consider any $i$. We have to show that $\pga \in \Phi_i$.%
	\footnote{Assume $(a_4, \arga, a) \nileadsto \pga$ (thus, $a$ attacks $\arga$). 
	By tf, $(a_4, \arga, a, \ar_2, …) \ileadsto \pga$.
	We get $(a_3, \arga, a, \ar_2) \ileadsto \pga$ by answ.
	By reinst: $(\ar_2)$ YES.}
